Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2014-01-06 paper/ms.tex
revised conclusions

In this \documentname we look for constraints on the preferred mass
and occupation fraction of dark matter halos hosting Lyman Alpha Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We perform this study
paying special attention to the impact of cosmic variance on these
results. To this end we build a large number of mock catalogs matching
observational geometries. The mocks are constructed from a N-body simulation
following a simple recipe to assign a single LAE to each halo. Only
a fraction $f_{\rm occ}$ of halos within a mass range  $M_{\rm
  min}<M_{\rm h}<M_{\rm   max}$ can host a LAE.

We then proceed with a
thorough exploration of the space of free parameters $M_{\rm min},
M_{\rm max}$ and $f_{\rm occ}$ to find mocks that are consistent with
two observational constraints: the surface number density and the
angular correlation function. Out of the initial $9000$ 
combinations of parameters in the model we find $40$ combination of parameters
consistent with observations.

{\bf We find that the information for the number density constrains
  the minium mass to the range
  $10^{10}\hMsun < M_{\rm min} < 10^{11.5}\hMsun$ while the clustering
  constrains the upper mass $M_{\rm    max} < 10^{12}\hMsun$. However,
  this is insufficient to impose a tight constraint  on the escape
  fraction.} Furthermore, we find that the mass range defined as
$\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm min}$ is always
$\Delta M\leq 1.0$.

{\bf Our approach has allowed the discovery of possible models that have
been overloooked so far in the literature. A central role in these new
models is played by the wide range of escape fraction values. A wide
range in the escape fraction is unexpected from previous studies that
unanimously favor low values for the occupation
fractions \citep[i.e.][]{Gawiser2007,Ouchi2010}. The larges
innnovation in our analysis is the construction of mock obsevations
that directly explore the impact of cosmic variance, which seems to be
at the root of the large uncertainties in the escape fraction.}

Nevertheless, all the halo mass ranges deduced previous analysis are a
subset of the models we find in this \documentname.  All of our
preferred models support the notion that the most massive halos at
$z=3.1$ do not host detectable LAEs. The existence of a narrow mass
range for halos hosting LAEs, $\Delta  M\leq 1.0$ dex, has two
different explanations at least. The first is having LAEs with a
decreasing \ly escape fraction with increasing mass, the second is
having larger screening effects by neutral Hydrogen around the most
massive systems \citep{Laursen2009,ForeroRomero2011}.

We also find extreme cases with a very narrow $\Delta M<0.3$, meaning that
there is barely a factor of $\sim 2$ between the minimum and halo
mass. If these models can be confirmed, it would mean a great
challenge for galaxy formation models to explain how is that LAEs can
be hosted in such a narrow range of halo mass.

In summary, we show that {\bf the currently available spatial
  information of LAEs cannot put a tight constraint on their
  occupation fraction of dark matter halos}. This information can
confidently be used to establish that these galaxies are hosted in low
mass halos with masses within a range spanning less than a factor of
ten. {\bf From the point of view of Dark Energy surveys such as
  HETDEX, the narrow mass range for LAEs can be seen as an advantage,
  due to its simplicity, for their modeling aiming at recovering
  cosmological parameters.} We foresee that the new observations with
new instruments (such as MUSE, Hyper SuprimeCam and HETDEX) covering
larger fields and a wider range of luminosities will be key in reducing cosmic variance and imposing tighter constraints on
the properties of dark matter halos hosting LAEs. At the same time,
additional modeling for \ly radiation transfer is needed to put a
tighter constrain on the \ly escape fraction in high redshift
galaxies, also paying attention to newly framed physical phenomena,
such as the stochasticity \citep{ForeroRomero2013} in the star
formation process, which might play a role in inducing detection
biases in high redshift LAEs.

>>>>

In this \documentname we look for constraints on the preferred mass
and occupation fraction of dark matter halos hosting Lyman $\alpha$ Emitters at
redshift $z=3.1$ in a $\Lambda$CDM cosmology. We perform this study
paying special attention to the impact of cosmic variance on these
results. To this end we build a large number of mock catalogs matching
observational geometries. The mocks are constructed from a N-body simulation
following a simple recipe to assign a single LAE to each halo whereby only
a fraction $f_{\rm occ}$ of halos within a mass range  $M_{\rm
  min}\leq M_{\rm h}\leq M_{\rm   max}$ can host a LAE.

We then proceed with a
thorough exploration of the space of free parameters $M_{\rm min},
M_{\rm max}$ and $f_{\rm occ}$ to find mocks that are consistent with
two observational constraints: the surface number density distribution
and the angular correlation function. Out of the initial $9000$ 
combinations of parameters in the model we find $40$ combination of parameters
consistent with observations.

{\bf We find that the information for the number density constrains
  the minium mass to the range
  $10^{10}\hMsun < M_{\rm min} < 10^{11.5}\hMsun$ while the clustering
  constrains the upper mass $M_{\rm    max} < 10^{12}\hMsun$. However,
  this information insufficient to impose a tight constraint  on the escape
  fraction. Additionaly, we find that the mass range defined as
$\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm min}$ is $\Delta
M\leq 1.0$, with a large fraction of models with $\Delta M\leq 0.5$.}

{\bf Our approach has allowed the discovery of possible models that have
been overloooked so far in the literature. A central role in these new
models is played by the wide range of escape fraction values,
something unexpected from previous studies \citep[i.e.][]{Gawiser2007,Ouchi2010}  that unanimously favor low values on the range of $10^{-2}$. The largest
innovation in our analysis is the construction of mock obsevations
that directly explore the impact of cosmic variance.}

Nevertheless, all the halo mass ranges deduced previous analysis are a
subset of the models we find in this paper.  All of our
preferred models support the notion that the most massive halos at
$z=3.1$ do not host detectable LAEs. {\bf Furthermore, we claim that the
mass range for halos hosting LAEs is narrow, $\Delta M\leq 1.0$, with
a majority of models with $\Delta M\leq 0.5$.}

The existence of models with narrow mass ranges has at least two
different explanations. The first is having LAEs with a decreasing \ly
escape fraction with increasing mass, the second is having larger
screening effects by neutral Hydrogen around the most massive systems
\citep{Laursen2009,ForeroRomero2011}.

We also find extreme cases with a very narrow $\Delta M<0.3$, meaning that
there is barely a factor of $\sim 2$ between the minimum and halo
mass. If this can be confirmed, it would mean a great
challenge for galaxy formation models to explain how is that LAEs can
be hosted in such a narrow range of halo mass.

In summary, we show that {\bf the currently available spatial
  information of LAEs cannot put a tight constraint on their
  occupation fraction of dark matter halos}. This information can
confidently be used to establish that these galaxies are hosted in low
mass halos with masses spanning less than one order of magnitude. {\bf
  From the point of view of Dark Energy surveys such as the Hobby-Eberly
  Telescope Dark Enery Experiment (HETDEX) the
  narrow mass range for LAEs can be seen as an advantage because its
  simplicity in their modeling aiming at recoverin cosmological
  parameters.} We foresee that the observations with new instruments
(such as the Multi Unit Spectroscopic Explorer (MUSE), Hyper
SuprimeCam and HETDEX) covering larger fields and a wider range of
luminosities will be key in reducing cosmic variance and imposing
tighter constraints on the properties of dark matter halos hosting
LAEs. At the same time, additional modeling for \ly radiation transfer
is needed to put tighter constraints on  occupation fraction in high
redshift halos, also paying attention to other physical phenomena,
such as the stochasticity \citep{ForeroRomero2013} in the star
formation process, which might play a role in inducing detection
biases in high redshift LAEs.
